### 2022-06-06 paper/localpsf.tex
edits to paper based on discussion with Noemi and Tucker

\paragraph{(2) Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank (maximum rank among the blocks of the H-matrix). To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulses used in the RBF interpolation. This is done using a precomputed KD tree of sample points, and requires $O(\numnbr \log \nsamplepts)$ floating point and logical elementary operations. Second, one must find the mesh cells that the points $\{z_i\}_{i=1}^{\numnbr}$ reside. This is done using an AABB tree of mesh cells, and requires $O(\numnbr \log \fedim)$ elementary operations. Third, we must evaluate finite element basis functions on each of those cells, which requires $O(\numnbr)$ elementary operations. Finally, we must solve a $\numnbr \times \numnbr$ linear system to perform the radial basis function interpolation, which requires $O(\numnbr^3)$ elementary operations, which typically is negligible. Therefore, building the H-matrix requires
\begin{equation*}
	O\left(\left(\hrank^2 \fedim \log \fedim\right)\left(\numnbr \log \fedim + \numnbr^3\right)\right) \qquad \text{elementary operations}.
\end{equation*}
This follows from the above discussion, and the fact that $\fedim \ge \nsamplepts$.

\paragraph{(3) Performing linear algebra operations with the H-matrix} It is well-known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion require
\begin{equation*}
	O(\hrank^2 \fedim \log(\fedim)^a) \qquad \text{elementary operations},
\end{equation*}
where $a \in \{0,1,2,3\}$. The precise cost depends on the type of H-matrix used and the operation being performed; for more details see \cite{GrasedyckHackbusch03}. The rational method for modifying $\mathbf{A}_H$ to be positive semi-definite, described in Appendix \ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the order of the rational function used.

>>>>

\paragraph{(2) Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix. To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulses used in the RBF interpolation. This is done using a precomputed KD tree of sample points, and requires $O(\numnbr \log \nsamplepts)$ floating point and logical elementary operations. Second, one must find the mesh cells that the points $\{z_i\}_{i=1}^{\numnbr}$ reside. This is done using an AABB tree of mesh cells, and requires $O(\numnbr \log \fedim)$ elementary operations. Third, we must evaluate finite element basis functions on each of those cells, which requires $O(\numnbr)$ elementary operations. Finally, we must solve a $\numnbr \times \numnbr$ linear system to perform the radial basis function interpolation, which requires $O(\numnbr^3)$ elementary operations, which typically is negligible. Therefore, building the H-matrix requires
\begin{equation*}
	O\left(\left(\hrank^2 \fedim \log \fedim\right)\left(\numnbr \log \fedim + \numnbr^3\right)\right) \qquad \text{elementary operations}.
\end{equation*}
%This follows from the above discussion, and the fact that $\fedim \ge \nsamplepts$.

\paragraph{(3) Performing linear algebra operations with the H-matrix} It is well-known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion, and low rank updates require
\begin{equation}
\label{eq:H_matrix_cost_operations}
	O(\hrank^2 \fedim \log(\fedim)^a) \qquad \text{elementary operations},
\end{equation}
where $a \in \{0,1,2,3\}$ \cite{GrasedyckHackbusch03}. The precise cost depends on the type of H-matrix used and the operation being performed. In this paper, we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian. Then, we use one matrix inversion to form the inverse of the Hessian, which is used as a preconditioner for the Newton linear system. We perform low-rank updates to the H-matrix as the Newton iterations proceed, as per the discussion in Appendix \ref{sec:recycle_krylov}. The rational method for modifying $\mathbf{A}_H$ to be positive semi-definite, described in Appendix \ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the order of the rational function used.
